Value-Suppressing Uncertainty Maps

### 2017-03-30 paper/figures.tex
Add example figures
checkin

@@

>>>>

\newcommand{\conditionFig}{
	\begin{figure*}
		\centering
		\includegraphics[width=0.8\textwidth]{conditions.pdf}
		\caption{The three graph types in our evaluation. Bivariate color ramps ensure that there are orthogonal mappings for each combination of value and uncertainty, but, since color channels are not separable, can afford only a few discrete colors before color categories become unacceptably close, perceptually. Juxtaposed maps, by keeping these channels separate, afford a larger range of colors, but require a visual search task to recover both value and uncertainty in a region. VSUMs have many of the advantages of bivariate maps, but assign more color categories to certain data, at the expense of ambiguity when uncertainty is high.}
		\label{fig:conditions}
	\end{figure*}
}

@@

\newcommand{\airlineFig}{
\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{airline-vsum.png}
	\caption{Flight delay data visualized with VSUM.}
	\label{fig:airlineVsum}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{airline-2d.png}
	\caption{Flight delay data visualized with 2D uncertainty map.}
	\label{fig:airline2d}
\end{figure}
}

\newcommand{\viralFig}{
\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{viral-vsum.png}
	\caption{Virus genome data for reads between position $1309$ and $10124$ visualized with VSUM.}
	\label{fig:viralVsum}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{viral-2d.png}
	\caption{Virus genome data visualized with 2D uncertainty map.}
	\label{fig:viral2d}
\end{figure}
}

### 2017-03-30 paper/template.tex
checkin

We performed a crowdsourced experiment on Amazon's Mechanical Turk in order to evaluate the effectiveness of VSUMs for integrating uncertainty and value information in thematic maps. This focus on integration meant that we limited our experimental tasks to scenarios where the participants needed to consider both value and uncertainty before making a decision. After a pretest and short tutorial, we gave participants two main tasks:

\begin{enumerate}
	\item [T1] an \textbf{identification} task, where we gaves participants charts with value and uncertainty information, and asked them to locate specific regions. E.g., ``click on the region of the chart with highest uncertainty \emph{and} lowest value.''
	\item [T2] a \textbf{prediction} task, similar to Battleship, where we gave participants charts with both \emph{forecast} and \emph{forecast uncertainty} information, and asked the participants to place tokens on the board in order to optimize expected value. E.g., ``place your $4$ ships on the safest locations on the board.''
\end{enumerate}
We discuss the two experimental tasks in more detail in the following sections.

After completion of the main task, we solicited demographic information, and performed a domain-agnostic general risk aversion assay~\cite{mandrik2005exploring}.

@@

@@

For the prediction task, participants were given the rules of a game similar to Battleship. Greis et al.~\cite{greis2016decision} employ these game-like experimental tasks to frame uncertainty information, which can be abstract or complex, to the general audience. In our task, the participant and a (fictional) adversary have to place tokens representing ships on a spatial grid. Without direct access to the location of the opponent's ships, the player then has to place tokens representing missile strikes on their opponent's grid. If a player fires a missile at a square containing a ship, then that ship is sunk. The objective is the maximize the number of enemy ships hit, while minimizing the number of your own ships that are hit. In our task, participants were given a chart encoding predictions on the location of enemy ships (in the attacking case) or potential missile strikes (in the defending case). The \emph{value} component of the prediction was the likelihood of a item being present in a particular region. The \emph{uncertainty} component was the confidence in this prediction.

@@

@@

We limited our population to Turkers from the United States, with a prior task approval rate of at least 90\%. As the experimental tasks required multi-hue color perception, we presented participants with a set of Ishihara plates as a pretest, and excluded participants who either misidentified the values in the plates, or who self-reported as having a color vision deficiency (CVD) in the post-test. Based on piloting, we paid participants \$X dollars an hour, for a target rate of \$8/hour.

>>>>

\conditionFig

We performed a crowdsourced experiment on Amazon's Mechanical Turk in order to evaluate the effectiveness of VSUMs for integrating uncertainty and value information in thematic maps. This focus on integration meant that we limited our experimental tasks to scenarios where the participants needed to consider both value and uncertainty before making a decision. We gave participants two main tasks:

\begin{enumerate}
	\item [T1] an \textbf{identification} task, where we gaves participants charts with value and uncertainty information, and asked them to locate specific regions. E.g., ``click on the region of the chart with highest uncertainty \emph{and} lowest value.''
	\item [T2] a \textbf{prediction} task, similar to Battleship, where we gave participants charts with both \emph{forecast} and \emph{forecast uncertainty} information, and asked the participants to place tokens on the board in order to optimize expected value. E.g., ``place your $4$ ships on the safest locations on the board.''
\end{enumerate}

We discuss the two experimental tasks in more detail in the following sections.

For each task, participants saw one of three types of thematic maps. Either two \textbf{juxtaposed} maps side-by-side, with separate encodings for value and uncertainty, a traditional \textbf{bivariate} map, where value and uncertainty are orthogonally encoded, or a \textbf{VSUM}. Figure \ref{fig:stimuli} shows examples of these factor levels.

@@

For the identification task, we gave participants bivariate representations of uncertainty, and asked them in a written prompt to identify a location meeting a given criteria, e.g., ``Click on the location with the greatest uncertainty.'' There were three potential criteria, mapping to the three vertices of a VSUM: 1) high value, low uncertainty, 2) low value, low uncertainty, and 3) high uncertainty. In each map,

@@

For the prediction task, we gave participants the rules of a game similar to Battleship. Greis et al.~\cite{greis2016decision} employ these game-like experimental tasks to frame uncertainty information, which can be abstract or complex, to the general audience. In our task, the participant and a (fictional) adversary have to place tokens representing ships on a spatial grid. Without direct access to the location of the opponent's ships, the player then has to place tokens representing missile strikes on their opponent's grid. If a player fires a missile at a square containing a ship, then that ship is sunk. The objective is the maximize the number of enemy ships hit, while minimizing the number of your own ships that are hit. In our task, participants were given a chart encoding predictions on the location of enemy ships (in the attacking case) or potential missile strikes (in the defending case). The \emph{value} component of the prediction was the likelihood of a item being present in a particular region. The \emph{uncertainty} component was the confidence in this prediction.

@@

To avoid potential (in this case, incorrect) transfer learning, the attacking and defending used different heatmaps than the identification tasks, and from each other.

@@

We limited our population to Turkers from the United States, with a prior task approval rate of at least 90\%. As the experimental tasks required multi-hue color perception, we presented participants with a set of Ishihara plates as a pretest, and excluded participants who either misidentified the values in the plates, or who self-reported as having a color vision deficiency (CVD) in the post-test. Based on piloting, we paid participants \$2 dollars an hour, for a target rate of \$8/hour. After completion of the main tasks, we solicited demographic information, and performed a domain-agnostic general risk aversion assay~\cite{mandrik2005exploring}.
